Astrophysical calibration of gravitational-wave detectors

### 2015-09-22 docs/cosmic_cal.tex
A bit more edits and fixed the compile.

% acronym definitions
\acrodef{GW}[GW]{gravitational wave}
\acrodef{NS}[NS]{neutron star}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{MWEG}[MWEG]{Milky Way equivalent galaxies}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{NSBH}[NSBH]{neutron star-black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{GRB}[GRB]{gamma-ray burst}
\acrodef{sGRB}[sGRB]{short gamma-ray burst}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{EM}[EM]{electro-magnetic}
\acrodef{aLIGO}[aLIGO]{Advanced LIGO}
\acrodef{AdV}[AdV]{Advanced Virgo}
\acrodef{PSD}[PSD]{power spectral density}
\acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
\acrodef{ET}[ET]{Einstein Telescope}

@@

% what are sGRBs 
It is believed that \acp{sGRB} (those \acp{GRB} with duration ${<}2$ sec) are
emitted during the merger of
\acp{BNS}~\cite{1989Natur.340..126E},~\cite{1992ApJ...395L..83N}. The emmission
from these events is highly beamed along the binary rotation axis and hence
only potentially observable for a fraction of \ac{BNS} mergers. The distances
to these objects are obtained via identification of

The observed redshift range is from about 0.2 to 2 for short GRBs (sGRBs), with
a mean of about 0.4. The XRT commonly observes sGRBs up to z ≃ 0.5 but in 10.5
yr of operation Swift has not seen one with a measured redshift z < 0.1, i.e.,
the NS-NS merger aLIGO target GW horizon ∼400 Mpc, suggesting that such events
may be quite rare.
Estimates for jet beaming are θj ∼ 5◦ for lGRBs and θj ∼ 5 − 15◦ for sGRBs
(Burrows et al. 2006, Grupe et al. 2006, Fong et al. 2012). Beaming angles for
sGRBs are still highly uncertain. The beaming factors fb = 1 − cos θj ≃ θj2/2
are roughly 1/300 for lGRBs and 1/30 for sGRBs. Based on the observed rate of
sGRBs by Swift, Coward et al. (2012) estimate a LVC detection rateof∼3−30yr−1
forθj ≃ 15◦. Chen&Holz(2013)claim3−7yr−1 forGRBGW+EM detections. Kelley et al.
(2013) estimate the rate of Swift or Fermi observations joint with LVC
detections to be ∼0.07 yr−1. Siellez, Boe ̈r, & Gendre (2014) consider current
and future high energy missions and estimate a rate of simultaneous GW+EM
detections of ∼0.1 − 4 yr−1 in the LVC era. Wanderman & Piran (2015) estimate a
co-detection rate LVC+Fermi of 0.1 − 1 yr−1 and LVC+Swift of 0.02 − 0.14 yr−1.

>>>>

% acronym definitions
\acrodef{GW}[GW]{gravitational wave}
\acrodef{NS}[NS]{neutron star}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{MWEG}[MWEG]{Milky Way equivalent galaxies}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{NSBH}[NSBH]{neutron star-black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{GRB}[GRB]{gamma-ray burst}
\acrodef{sGRB}[sGRB]{short gamma-ray burst}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{EM}[EM]{electro-magnetic}
\acrodef{aLIGO}[aLIGO]{Advanced LIGO}
\acrodef{AdV}[AdV]{Advanced Virgo}
\acrodef{PSD}[PSD]{power spectral density}
\acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
\acrodef{ET}[ET]{Einstein Telescope}
\acrodef{XRT}[XRT]{x-ray telescope}

@@

% what are sGRBs 
It is believed that \acp{sGRB} (those \acp{GRB} with duration ${<}2$ sec) are
emitted during the merger of
\acp{BNS}~\cite{1989Natur.340..126E},~\cite{1992ApJ...395L..83N}. The emmission
from these events is highly beamed along the binary rotation axis and hence
only potentially observable for a fraction of \ac{BNS} mergers. If the event
exhibits an optical afterglow then the host galaxy can be identified from which
a redshift can be obtained.  The fraction of events with associated redshifts
is ${\sim}1/3$. The range to which the \emph{Swift} \ac{XRT} has detected
\acp{sGRB} is $z{\sim}2$ and the nearest event with associated redshift is at
$z{sim}0.1$~\cite{2015GCN..17278...1C} which is approximately equal to the
horizon range of the advanced \ac{GW} detector network.

% this is copied text for reference !! 
%Estimates for jet beaming are θj ∼ 5◦ for lGRBs and θj ∼ 5 − 15◦ for sGRBs
%(Burrows et al. 2006, Grupe et al. 2006, Fong et al. 2012). Beaming angles for
%sGRBs are still highly uncertain. The beaming factors fb = 1 − cos θj ≃ θj2/2
%are roughly 1/300 for lGRBs and 1/30 for sGRBs. Based on the observed rate of
%sGRBs by Swift, Coward et al. (2012) estimate a LVC detection rateof∼3−30yr−1
%forθj ≃ 15◦. Chen&Holz(2013)claim3−7yr−1 forGRBGW+EM detections. Kelley et al.
%(2013) estimate the rate of Swift or Fermi observations joint with LVC
%detections to be ∼0.07 yr−1. Siellez, Boe ̈r, & Gendre (2014) consider current
%and future high energy missions and estimate a rate of simultaneous GW+EM
%detections of ∼0.1 − 4 yr−1 in the LVC era. Wanderman & Piran (2015) estimate a
%co-detection rate LVC+Fermi of 0.1 − 1 yr−1 and LVC+Swift of 0.02 − 0.14 yr−1.
